---
title: "iLab 01 - EDA"
author: "Nathaniel Evans"
date: "October 1, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tidyverse) # all the good stuff
library(readxl) # for reading in xlsx files
library(here) # for here!
library(janitor) # for clean_names & tabyl
library(knitr) # for kable

library(reprohealth)
data <- wb_reprohealth %>% mutate(year=parse_number(year))
head(data)
```
## **reprohealth EDA**
###How many variables/columns?
There are 9 variables of types character (3) and numeric (6)
```{r}
summary(data)
```
##How many rows/observations?
There are 3296 rows in this dataset. 
```{r}
n_rows <- data %>% count()
print(n_rows)
```
##Which variables are numbers?
The following variables are numeric of type <dbl> : 
*year
*edu_gen_parity
*gdp_per_cap
*life_exp
*mat_mortality
*tot_fertility
```{r}
data %>% glimpse()
```
##Which are categorical variables (numeric or character variables with variables that have a fixed and known set of possible values; aka factor variables)?
The three character types and year have set possible values: 
*year (16) [2000-2015]- technically of type <dbl> but since it's a year, they must be ints within bounded possible years (eg no years before XX and no years after 2018 -today!) so it _can_ be treated as a factor. Not sure if you'd want to or not though. 
*country (206)
*continent (5)
*region (22)
```{r}
yrs <- data %>% distinct(year) %>% count()
cntry <- data %>% distinct(country) %>% count()
cntnt <- data %>% distinct(continent) %>% count()
rgn <- data %>% distinct(region) %>% count()

print(c("distinct years: ", yrs))
print(c("distinct countries: ", cntry))
print(c("distinct continents: ", cntnt))
print(c("distinct regions: ", rgn))

data$year %>% summary()

```
Complete this sentence: "There is one row per."
```{r}
head(data)
```

