---
title: "iLab 01 - EDA"
author: "Nathaniel Evans"
date: "October 1, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tidyverse) # all the good stuff
library(readxl) # for reading in xlsx files
library(here) # for here!
library(janitor) # for clean_names & tabyl
library(knitr) # for kable

library(reprohealth)
data <- wb_reprohealth %>% mutate(year=parse_number(year))

head(data)
```

## **reprohealth EDA**
###How many variables/columns?
There are 9 variables of types character (3) and numeric (6)  
```{r}
summary(data)
```
##How many rows/observations?
There are 3296 rows in this dataset.   
```{r}
n_rows <- data %>% count()
print(n_rows)
```
##Which variables are numbers?
The following <span style="color:deeppink"> 6 </span> variables are numeric of type <dbl> :   
* year  
* edu_gen_parity  
* gdp_per_cap  
* life_exp  
* mat_mortality  
* tot_fertility  
```{r}
data %>% glimpse()
```
##Which are categorical variables (numeric or character variables with variables that have a fixed and known set of possible values; aka factor variables)?
The three character types and year have set possible values:  
* year (16) [2000-2015]   
  + technically of type <dbl> but since it's a year, they must be ints within bounded   possible years (eg no years before XX and no years after 2018 -today!) so it _can_ be treated as a factor. Not sure if you'd want to or not though.     
* country (206)  
* continent (5)  
* region (22)  
```{r}
yrs <- data %>% distinct(year) %>% count()
cntry <- data %>% distinct(country) %>% count()
cntnt <- data %>% distinct(continent) %>% count()
rgn <- data %>% distinct(region) %>% count()

print(c("distinct years: ", yrs))
print(c("distinct countries: ", cntry))
print(c("distinct continents: ", cntnt))
print(c("distinct regions: ", rgn))

data$year %>% summary()

```
Complete this sentence: "There is one row per." <span style="color:deeppink"> **observation of country fertility related metrics over time** </span>
```{r}
head(data)
```

# Part 2 -------------------------------
## Gapminder + new data source

```{r}
setwd("C:/Users/natha/Documents")
alc <- read_xlsx(("indicator alcohol consumption  20100830.xlsx"),
                      sheet = 1) %>% 
                      clean_names() # highly recommended

alc2 <- alc %>%
  rename(country = 1) %>% 
  gather(year, x1, -country) %>% 
  mutate(year = readr::parse_number(year)) %>% 
  filter(!is.na(x1))

glimpse(alc2)
```

###How many variables/columns?  
3: chr(1) and dbl(2)  

###How many rows/observations? 
411 observations (after dropping na's)  

###Which variables are numbers?  
year and x1 (alc consumption (Liters)) 

###Which are categorical variables (numeric or character variables with variables that have a fixed and known set of possible values; aka factor variables)?   
Just country. 
   
###Complete this sentence: "There is one row per." 
*observation* or specifically *observation characterizing year, country and avg consumption of alcohol in pop >= 15 yo*

## joining the data (tables)
```{r}
data_jnd <- left_join(data, alc2) 
glimpse(data_jnd)
```
#Part 3 ----------------------
### "Your goal during EDA is to develop an understanding of your data. The easiest way to do this is to use questions as tools to guide your investigation.EDA is fundamentally a creative process. And like most creative processes, the key to asking quality questions is to generate a large quantity of questions." 


<span style="color:deeppink"> I'm going to add a little bit here to explore the data, initially, like the example does: Visualize Max/Mins and distribution by continent. This violin plot is my new favorite. Especially overlaying a point distribution of the data on top.


<span style="color:deeppink"> From this graph, we can see that while europe is by far the highest on average, it has a few countries that don't drink much at all. Additionally, while asia is predominantly non-drinking, there are a few that compete with even the highest of europe and has the most significant range of values. Oceania, has a fairly odd distribution that looks like two separate populations and may be described by the indonesian countries that are more similar to asia and the "western" countries (Australia, NZ, Tazmania) Needs some more exploration to say for sure. Americas have a pretty interestingly tight distribution and Africa has a noticeably significant range of values. 

</span>
```{r}
data_jnd %>% filter(!is.na(x1)) %>% group_by(continent) %>% ggplot(aes(x = continent, y = x1, color=continent )) + geom_violin() + geom_point()
```


#task 1
##How does [alcohol consumption] (x1) vary across different continents?
```{r}
# what years does this include? 
data_jnd %>% filter(!is.na(x1), !is.na(gdp_per_cap)) %>% select(year) %>% summary() 
# lets create median of the years available 
data3 <- data_jnd %>% filter(!is.na(x1), !is.na(gdp_per_cap)) %>% group_by(continent) %>% summarize(med_x1 = median(x1), med_gdp = median(gdp_per_cap)) 
# get rid of med_gdp, we'll use that next. Generate table
dg <- data3 %>% select(-med_gdp)
dg

#plot a bar plot 
dg %>% ggplot(aes(x=continent, y=med_x1, fill= continent)) + geom_col() + ggtitle("Alcohol consumption by continent")
```

From this table and chart, we can show that between the years of 2000 and 2008, the median liters of alcohol consumed per person per year varied significantly between continents. Europe was by far the leader in alcohol consumption, with americas being just under half as much. The remaining countries were all about half of the americas. 

## task 2 : how do countries alcohol consumption change over time? 
Find countries with interesting stories. Open-ended and, therefore, hard. Promising but unsuccessful attempts are encouraged. This will generate interesting questions to follow up on in class.

```{r}
data4 <-  data_jnd %>% filter(!is.na(x1), !is.na(gdp_per_cap)) %>% group_by(country)
data4 %>% ggplot(aes(x=year, y=x1, color=country)) + geom_line() + ggtitle("country alcohol consumption over time") + theme(legend.position = "none")
```

Whoops! But very interesting, we only have one country with data from 2000 to 2008, the rest only span from 2005 - 2008. I'll filter to only include 2005-2008. Also, there are way too many countries to compare this way, so we'll filter to just Africa. Lastly, there is very little change over the years in countries with x1 less than 5 so we'll only include countries with x1 > 5. 

```{r}
head(data4)
data4 %>% filter(year>=2005, continent=='Africa', x1>5) %>% ggplot(aes(x=year, y=x1, color=country)) + geom_line() + ggtitle("country alcohol consumption over time")
```
This didn't come out quite as nice as I'd hoped but there are some notable countries, in particular, Uganda, that top pink line, had a significant increase in alcohol consumption. To dig into that a little more... 

```{r}
ug_data <- data4 %>% filter(country=="Uganda")
ug_data %>% glimpse() 
```

This table shows that the only other significant increase over these two observations were gdp_per_cap, which actually makes sense; I would expect that as gdp increases, so does alcohol consumption. 

##task 3 - does alc consumption correllate to gdp_per_cap? 
```{r}
data4 %>% ggplot(aes(x=x1, y=gdp_per_cap, color=year)) + geom_point() + ggtitle("alc consumption vs gdp/cap") +  geom_smooth(method='lm')
mod <- lm(formula=x1~gdp_per_cap, data=data4)
summary(mod)
```


It looks like there is weak correlation, but a residual analysis is definitely needed and I 'd guess that there really isn't much predictive value in this model. However, it does seem like there may be two different groupings of this data that follow more predicatble trends, simply, data above the fitted line and data below the fitted line might have better predictive succuess if they could be separated by another variable. 

```{r}
data4 %>% ggplot(aes(x=x1, y=gdp_per_cap, color=continent)) + geom_point() + ggtitle("alc consumption vs gdp/cap (by continent)") 

# remove europe
data4 %>% filter(continent != "Europe", continent != "Asia") %>% ggplot(aes(x=x1, y=gdp_per_cap)) + geom_point() + ggtitle("alc consumption vs gdp/cap in americas, africa, oceania") +  geom_smooth(method='lm') + scale_y_log10()
```

Coloring the plot by continent shows that a significant portion of the outlying data is from Europe and Asia, and by removing those, it *looks* like there is increased correllation, and more importantly a linear model would have more predictive value. Additionally, by scaling gdp_per_cap by log10, it creates a better distribution of values and shows the trend. There are a great number of reasons that could account for this, I'd hazard that the most likely reasons stem from culture and/or production.  


Credit to Justine and Alfonso for explaining how to get rmarkdown to add line breaks. I was going nuts trying to figure that out. 


