---
title: "iLab3: Comparing Means"
author: "Nathaniel Evans"
date: "November 12, 2018"
output:
  html_document: 
    theme: flatly 
    highlight: kate
    toc: true
    toc_depth: 2
    toc_float : true
    smooth_scroll: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(infer)
library(skimr)
library(broom)
library(GGally)
```


#Overview 

<span style="color:deeppink"> Minor changes in this color </span>

## Data: [csv](https://raw.githubusercontent.com/apreshill/ohsu-basic-stats/master/data/fishermen_mercury.csv)

## Paper: [Science Direct](https://www.sciencedirect.com/science/article/pii/S0269749199002614?via%3Dihub#TBL2)

## Codebook: 

###Variables in this dataset include: 
####Fisherman indicator (fisherman)
####Age in years (age)
####Residence Time in years (restime)
####Height in cm (height)
####Weight in kg (weight)
####Fish meals per week (fishmlwk)
####Parts of fish consumed:0=none, 1=muscle tissue only, 2=mt and sometimes whole fish, 3=whole fish (fishpart)
####Methyl Mercury in mg/g (methyl_mercury)
####Total Mercury in mg/g (total_mercury)

##Research Question
###**Do fishermen have different levels of total mercury than non-fishermen?**

# EDA 
#### Honestly, I did a significant amount of EDA the last time using this dataset, so I'm going to leave filling out the EDA to the revision portion. 
## Download, import data
#### $curl https://raw.githubusercontent.com/apreshill/ohsu-basic-stats/master/data/fishermen_mercury.csv >> fisherman_data.csv

```{r}
mercury <- read_csv("fisherman_data.csv",
                    col_types = cols(
                      fisherman = col_factor(levels = NULL))
                    )
```

##Looking at the raw values.
```{r}
mercury %>% glimpse()

```


##Computing summary statistics of the variables of interest.
```{r}
mercury %>% skim()

summary(mercury)

mercury %>% filter(fisherman==1) %>% summary()

mercury %>% filter(fisherman==0) %>% summary()
```


##Creating informative visualizations.

```{r, message=FALSE}
ggpairs(mercury)

```

# Check Conditions</br>  

####Remember that in order to use the short-cut (formula-based, theoretical) approach, we need to check that some conditions are met. The infer package does not automatically check conditions for the theoretical methods to work. In order for the results of the t-test to be valid, three conditions must be met:

##Independent observations in both samples </br>     
<span, color='red'> Each observation is independant of one another. An example of dependant observations would be cortisol levels before and after pregnancy. </span>
##Nearly normal populations OR large sample sizes (n???30)  </br>
<span, color='red'> We have large sample sizes: 100 fisherman, 35 non-fisherman</span>
##Independently selected samples
<span, color='red'> Two know that our observations are independant in both populations, we need to ask the question, how was the data curated? If the data was self-reported, misrepresentative location sampling or misrepresentative temporal sampling then there may be biases that make our population's observations dependant on eachother. For this dataset, since we know nothing about the data aquisiton methods, we will assume that the observations are independant </span>  
##For this lab, we'll also assume that the variance in the response/outcome variable is the same for both groups </br>
<span, color='red'> What range of variance is considered 'the same' ? </span>


## t-test for two independant samples 

###Z-Score Formula 
$$Z = \frac{x-\mu}{\sigma} $$

### T-statisitc Formula 
$$ T = \frac{ (\bar{x_1} - \bar{x_2}) - (\mu_1 - \mu_2) }{s_p \times \sqrt{\frac{1}{n_1} + \frac{1}{n_2}}} $$

$$ S_p = \sqrt{\frac{(n_1-1) \times s_1^2 + (n_2-1) \times s_2^2} {n_1 + n_2 -2}} $$

###"A t-value of 0 indicates that the sample results exactly equal the null hypothesis. As the difference between the sample data and the null hypothesis increases, the absolute value of the t-value increases."

### BUT under null hypthesis mu1-mu2 = 0

## Test Statistic
```{r}
# n1 is non-fisherman, n2 is fisherman 
n1 <- mercury %>% filter(fisherman==0) %>% nrow()
n2 <- mercury %>% filter(fisherman==1) %>% nrow()
s1 <- (mercury %>% filter(fisherman==0))$total_mercury %>% sd()
s2 <- (mercury %>% filter(fisherman==1))$total_mercury %>% sd()

Sp = sqrt( ((n1-1)*s1^2 + (n2-1)*s2^2)/(n1+n2-2))

xbar1 <- (mercury %>% filter(fisherman==0))$total_mercury %>% mean()
xbar2 <- (mercury %>% filter(fisherman==1))$total_mercury %>% mean()

T = (xbar1 - xbar2) / (Sp * sqrt( (1/n1) + (1/n2) ))

Sp 
T 
```
### Why is my Test Statistic negative? did I misinterpret xbar and mu? 

<span style="color:deeppink"> Fixed my upper tcrit value </span>

```{r}
upper_tcrit <- qt(1-0.025, 133) # TWO TAIL; 0.05 / 2
test_stat<- mercury %>% 
  specify(total_mercury ~ fisherman) %>% 
  hypothesize(null = "independence") %>% 
  calculate(stat = "t", order = c(1,0))
#test_stat

test_stat %>%  visualize(method = "theoretical", 
            obs_stat = upper_tcrit, 
            direction = "both") # gives us shading
```

## Observed Effect 
```{r}
ttest <- mercury %>% 
  t_test(total_mercury ~ fisherman, order = c(1, 0), var.equal = TRUE)
ttest

t_test_CI <- c(ttest$lower_ci, ttest$upper_ci)
t_test_CI

df <- 10
ggplot(data.frame(x = c(-4, 4)), aes(x)) + 
  stat_function(fun = dt, args = list(df = df)) + # t dist
  stat_function(fun = dnorm, lty = 5, color = "red") # normal dist in red
```

##The theoretical p-value   

But, how can we calculate the p-value from this? Well the easy way is with your infer output :) But! Remember that your observed t-statistic is a quantile value for a statistic that we assume follows a t-distribution. How do we calculate the probability of getting a t-statistic as or more extreme than the one we got? We use pt()!

Use pt() to calculate the 2-sided p-value in R. Does it match the output from infer?

```{r}

mu_f <- (mercury %>% filter(fisherman==1))$total_mercury %>% mean()
#mu_f

t2 <- (mean(mercury$total_mercury)-mu_f)/(sd(mercury$total_mercury)/sqrt(length(mercury$total_mercury))) #this is compared to the entire population... but I think it's supposed to be just the non-fisherman
#t2

dof <- mercury %>% filter(fisherman==0) %>% nrow()  
#dof

# from above
t <- 2.777389	
# what about ncp? 

p1 <- pt(q=t, df=135-2) # this ends up being supper high
p2 <- 1-pt(q=t, df=135-2)

#p1
MY.P.VAL.FINALLY <- 2*p2

MY.P.VAL.FINALLY
```

<span style="color:deeppink"> Fixed my theoretical p value, used the wrong degree of freedom and end of the tail </span>

#### Well this is clearly incorrect but I'm not sure how to think about this conceptually to remedy it. (Not anymore)


##The theorectical confidence interval

Unfortunately, we currently can't get the confidence interval for theoretical methods using infer. But we can use formulas (yay):

$$ \bar{x_1} - \bar{x_2} \pm t * SE_{\bar{x1}-\bar{x_2}} $$

$$SE =  s_p \times \sqrt{\frac{1}{n_1} + \frac{1}{n_2}}$$ 

You should already have variables for the mean difference and the t-statistic critical value(s) (here, t???) from the top formula. In the bottom formula, that standard error should look familiar- it is the standard error of the difference between means. This was the denominator in your calculated t-statistic.  


Use these values to calculate the 95% confidence interval for the difference in means. Does it match the output from  infer? Does it contain zero? Will this interval ever not contain the observed difference in sample means?

```{r}
Tcrit <- qt(0.975, 133)
cUP <- abs(xbar1-xbar2) + Tcrit*(Sp * sqrt( (1/n1) + (1/n2) ) )
cLOW <- abs(xbar1-xbar2) - Tcrit*(Sp * sqrt( (1/n1) + (1/n2) ) )

theo_CI <- c(cLOW, cUP)

t_test_CI = c(cLOW, cUP)

cUP
cLOW
```
### YUP! same CI 


##Compare t-test to permutation   

Using ModernDive 10.7 as an example, as well as our previous resampling lab, conduct a permutation test to evaluate whether there is a mean difference in total mercury between fishermen and non-fishermen. What do you conclude, and how do these results compare to those based on the classical hypothesis test? What, if any, assumptions of each method are comfortable or uncomfortable with?

```{r}
set.seed(0)
mercury_diff <- mercury %>% 
  specify(total_mercury ~ fisherman) %>%  
  calculate(stat = "diff in means", order = c(1, 0))


null_distn <- mercury %>% 
  specify(total_mercury ~ fisherman) %>% 
  hypothesize(null = "independence") %>% 
  generate(reps = 1000, type = "permute") %>% 
  calculate(stat = "diff in means", 
            order = c("0", "1"))

null_distn %>% 
  visualize(obs_stat = mercury_diff, 
            direction = "two_sided") # t_test_CI from above, it would be nice to plot the null distribution too

```
<span style="color:deeppink"> Added a mercury difference instead of the confidence interval (as I had before) </span>

##Compare t-test to linear regression

Use a simple linear regression to analyze total_mercury ~ fisherman. Look carefully at the output, and compare to the output of the classical t-test. What do you notice is the same?

```{r}
my_lm <- lm(data= mercury, total_mercury ~ fisherman) 
summary(my_lm)
```

### the p-value is the same as the t-statistic above. 

<span style="color:deeppink"> There were a few minor mistakes in the visualizations and one misconception on my theoretical p-value but in majority I matched the confidence interval and p-value calculations by manual comparison.  </span>

# References 

### [1] N.B. Al-Majed and M.R. Preston (2000). "Factors Influencing the Total Mercury and Methyl Mercury in the Hair of Fishermen in Kuwait," Environmental Pollution, Vol. 109, pp. 239-250.